Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-09 manuscript.Rmd
results edits

The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms). This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work [@FranchakScott2021], we compared two types of models. To assess the accuracy of body position classification for each recording session, we held back the last 25% of a session's proximal comparison data as the testing set. The testing set was never used as training data. A *group model* was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in training. In contrast, *individual models* were trained using the first 75% of a recording session, then tested against the last 25%. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification.

@@

Overall accuracy (Figure \@ref(fig:metrics)A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Median overall accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]` for group models and `r metrics_lookup["Overall Accuracy", "median_Individual"]` for individual models. Although individual model accuracy was greater than group model accuracy, group model accuracy approached the level of agreement found between two human coders (median = .92) indicating that group models could adequately replace human coders. Still, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was considerably lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`). Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.

### Cohen's Kappa 
Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)B), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with median group kappas (`r metrics_lookup["Kappa", "median_Group"]`) somewhat worse compared with median individual kappas (`r metrics_lookup["Kappa", "median_Individual"]`). Guidelines for interpreting kappa statistics [@LandisKoch1977] consider 0.81–1.00 "Almost Perfect,” 0.61–0.80 “Substantial,” 0.41–0.60 “Moderate,” 0.21– 0.40 “Fair,” and 0–0.20 “Slight to Poor”, indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

As in past work, all body positions were accurately classified even though performance varied somewhat between positions. As Table \@ref(tab:metrics-by-class) shows, median kappa statistics were strongest for prone (group = `r metrics_long_summary["Prone Kappa", "median_Group"]`, individual = `r metrics_long_summary["Prone Kappa", "median_Individual"]`) and supine (group = `r metrics_long_summary["Supine Kappa", "median_Group"]`, individual = `r metrics_long_summary["Supine Kappa", "median_Individual"]`). Sitting performance fell in the middle, and was considerably worse for group models than individual models (group = `r metrics_long_summary["Sitting Kappa", "median_Group"]`, individual = `r metrics_long_summary["Sitting Kappa", "median_Individual"]`). Held (group = `r metrics_long_summary["Held Kappa", "median_Group"]`, individual = `r metrics_long_summary["Held Kappa", "median_Individual"]`) and upright (group = `r metrics_long_summary["Upright Kappa", "median_Group"]`, individual = `r metrics_long_summary["Upright Kappa", "median_Individual"]`) performance was the least accurate, however, median performance was still within the "Substantial" range.

@@

Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another. Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?). High PPV means that we can be confident in the event label. Table \@ref(tab:metrics-by-class) shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. All values were sufficient (> .8). Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV across classes for both model types.

>>>>

The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms). This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*. To assess the accuracy of body position classification for each recording session, we held back the last 25% of a session's proximal comparison data as the testing set. The testing set was never used as training data. A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75% of a recording session, then tested against the last 25%. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification. 
<!-- GRAPHIC FOR TWO KINDS OF MODELS? -->

@@

Overall accuracy (Figure \@ref(fig:metrics)A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Overall accuracy for group models  (*M* = `r metrics_lookup["Overall Accuracy", "mean_Group"]`) was slightly lower than accuracy for individual models (*M* = `r metrics_lookup["Overall Accuracy", "mean_Individual"]`). Although overall accuracy from our unsupervised, in-home data collection did not match the near-perfect accuracy  (.95-.98) found in prior in-lab studies [@FranchakScott2021; @AiraksinenRasanen2020], both models approached the level of agreement found between two human coders (*M* = .905). Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model. Visual inspection of Figure \@ref(fig:metrics)A shows that accuracy values were heavily skewed, with most approaching perfect accuracy but a few participants with very poor accuracy. Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = `r metrics_lookup["Overall Accuracy", "median_Group"]`; individual median accuracy = `r metrics_lookup["Overall Accuracy", "median_Individual"]`. Indeed, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was considerably lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`). Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.

### Cohen's Kappa 
Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)B), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with group kappas (*M* = `r metrics_lookup["Kappa", "mean_Group"]`) somewhat worse compared with individual kappas (*M* = `r metrics_lookup["Kappa", "mean_Individual"]`). Guidelines for interpreting kappa statistics [@LandisKoch1977] consider 0.81–1.00 "Almost Perfect,” 0.61–0.80 “Substantial,” 0.41–0.60 “Moderate,” 0.21– 0.40 “Fair,” and 0–0.20 “Slight to Poor”, indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

As in past work [@FranchakScott2021; @AiraksinenRasanen2020], all body positions were accurately classified even though performance varied somewhat between positions. As Table \@ref(tab:metrics-by-class) shows, mean kappa statistics were strongest for prone (group *M* =  `r metrics_long_summary["Prone Kappa", "mean_Group"]`, individual *M* =  `r metrics_long_summary["Prone Kappa", "mean_Individual"]`) and supine (group *M* =  `r metrics_long_summary["Supine Kappa", "mean_Group"]`, individual *M* =  `r metrics_long_summary["Supine Kappa", "mean_Individual"]`). Sitting performance fell in the middle, and was considerably worse for group models than individual models (group *M* =  `r metrics_long_summary["Sitting Kappa", "mean_Group"]`, individual *M* =  `r metrics_long_summary["Sitting Kappa", "mean_Individual"]`). Held (group *M* =  `r metrics_long_summary["Held Kappa", "mean_Group"]`, individual *M* =  `r metrics_long_summary["Held Kappa", "mean_Individual"]`) and upright (group *M* =  `r metrics_long_summary["Upright Kappa", "mean_Group"]`, individual *M* =  `r metrics_long_summary["Upright Kappa", "mean_Individual"]`) performance was the least accurate, however, average performance was still within the "Substantial" range.

@@

Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another. Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?). High PPV means that we can be confident in the event label. Table \@ref(tab:metrics-by-class) shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. All values were sufficient (> .8). Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types. Table \@ref(tab:metrics-table) shows the overall sensitivity and PPV across classes.

### 2023-03-09 manuscript.tex
results edits

The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms). This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work (Franchak et al., 2021), we compared two types of models. To assess the accuracy of body position classification for each recording session, we held back the last 25\% of a session's proximal comparison data as the testing set. The testing set was never used as training data. A \emph{group model} was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in training. In contrast, \emph{individual models} were trained using the first 75\% of a recording session, then tested against the last 25\%. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \ref{fig:metrics}, Table \ref{tab:metrics-table}, and Table \ref{tab:metrics-by-class} summarize the performance of group and individual models using standard metrics for classification.

@@

Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Median overall accuracy was 0.89 for group models and 0.93 for individual models. Although individual model accuracy was greater than group model accuracy, group model accuracy approached the level of agreement found between two human coders (median = .92) indicating that group models could adequately replace human coders. Still, it is notable that the worst-case accuracy from group models (minimum = 0.50) was considerably lower than in individual models (minimum = 0.73). Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.

@@

Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with median group kappas (0.77) somewhat worse compared with median individual kappas (0.85). Guidelines for interpreting kappa statistics (Landis \& Koch, 1977) consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

As in past work, all body positions were accurately classified even though performance varied somewhat between positions. As Table \ref{tab:metrics-by-class} shows, median kappa statistics were strongest for prone (group = 0.968, individual = 0.942) and supine (group = 0.907, individual = 0.983). Sitting performance fell in the middle, and was considerably worse for group models than individual models (group = 0.816, individual = 0.915). Held (group = 0.732, individual = 0.826) and upright (group = 0.707, individual = 0.822) performance was the least accurate, however, median performance was still within the ``Substantial'' range.

@@

Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another. Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?). High PPV means that we can be confident in the event label. Table \ref{tab:metrics-by-class} shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. All values were sufficient (\textgreater{} .8). Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV across classes for both model types.

>>>>

The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms). This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work (Franchak et al., 2021), we compared two types of models: \emph{group models} and \emph{individual models}. To assess the accuracy of body position classification for each recording session, we held back the last 25\% of a session's proximal comparison data as the testing set. The testing set was never used as training data. A group model was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75\% of a recording session, then tested against the last 25\%. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \ref{fig:metrics}, Table \ref{tab:metrics-table}, and Table \ref{tab:metrics-by-class} summarize the performance of group and individual models using standard metrics for classification.

@@

Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Overall accuracy for group models (\emph{M} = 0.85) was slightly lower than accuracy for individual models (\emph{M} = 0.92). Although overall accuracy from our unsupervised, in-home data collection did not match the near-perfect accuracy (.95-.98) found in prior in-lab studies (Airaksinen et al., 2020; Franchak et al., 2021), both models approached the level of agreement found between two human coders (\emph{M} = .905). Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model. Visual inspection of Figure \ref{fig:metrics}A shows that accuracy values were heavily skewed, with most approaching perfect accuracy but a few participants with very poor accuracy. Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = 0.89; individual median accuracy = 0.93. Indeed, it is notable that the worst-case accuracy from group models (minimum = 0.50) was considerably lower than in individual models (minimum = 0.73). Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.

@@

Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}B), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with group kappas (\emph{M} = 0.75) somewhat worse compared with individual kappas (\emph{M} = 0.82). Guidelines for interpreting kappa statistics (Landis \& Koch, 1977) consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

As in past work (Airaksinen et al., 2020; Franchak et al., 2021), all body positions were accurately classified even though performance varied somewhat between positions. As Table \ref{tab:metrics-by-class} shows, median kappa statistics were strongest for prone (group \emph{M} = 0.860, individual \emph{M} = 0.841) and supine (group \emph{M} = 0.764, individual \emph{M} = 0.912). Sitting performance fell in the middle, and was considerably worse for group models than individual models (group \emph{M} = 0.702, individual \emph{M} = 0.887). Held (group \emph{M} = 0.726, individual \emph{M} = 0.727) and upright (group \emph{M} = 0.673, individual \emph{M} = 0.741) performance was the least accurate, however, median performance was still within the ``Substantial'' range.

@@

Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another. Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?). High PPV means that we can be confident in the event label. Table \ref{tab:metrics-by-class} shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. All values were sufficient (\textgreater{} .8). Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types. Table \ref{tab:metrics-table} shows the overall sensitivity and PPV across classes.
